---
title: "Cookbook"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cookbook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(gsm)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This vignette contains sample code showing how to use the Good Statistical Monitoring `{gsm}` package using sample data from [`{clindata}`](https://github.com/Gilead-BioStats/clindata). For more information on the `{gsm}` package see the [package homepage](https://gilead-biostats.github.io/gsm/).

# Setup and Installation

Run the following:

```{r eval = FALSE, include = TRUE}
## Install devtools
install.packages('devtools')

## Install and load sample raw data
devtools::install_github("Gilead-BioStats/clindata", ref = "main")
library(clindata)

## Install and load gsm
devtools::install_github("Gilead-BioStats/gsm", ref = "main")
library(gsm)
```

# Example 1 - Adverse Events Metric - Scripted

This example uses the standard {gsm} analysis workflows to creates site-level Adverse Event scripts. See the [Data Analysis Vignette](https://gilead-biostats.github.io/gsm/articles/DataAnalysis.html) for more detail.

- **Example 1.1** calculates the Site-level AE rates.
- **Example 1.2** adds a filter to include only Serious Adverse Events (SAEs) and implements pipes to run through the workflow.
- **Example 1.3** generates bar charts showing SAE rates and z-scores by study.
- **Example 1.4** generates a scatter plot with confidence bound for SAE rates.

```{r file = system.file("examples", "1_AdverseEventKRI.R", package = "gsm"), eval = FALSE, include = TRUE}

```

# Example 2 - Adverse Events Metrics - Workflow

This examples introduces YAML workflows to re-generate the same results as in **Example 1** via a reusable pipeline. See the [Data Model Vignette](https://gilead-biostats.github.io/gsm/articles/DataModel.html) for more detail.

- **Example 2.1** runs the AE KRI workflow.
- **Example 2.2** updates the metadata to run country-level metrics.
- **Example 2.3** adds a filtering step to the workflow to generate the SAE metric.

```{r file = system.file("examples", "2_AdverseEventWorkflow.R", package = "gsm"), eval = FALSE, include = TRUE}

```

# Example 3 - Study-Level Reporting Workflows 

This example extends the previous examples to generate charts and reports for multiple KRIs. See the [Data Reporting Vignette](https://gilead-biostats.github.io/gsm/articles/DataReporting.html) for more detail.

- **Example 3.1** steps through several workflows to generate a report for all 12 standard site-level KRIs.
- **Example 3.2** automates data ingestion using `Ingest()` and `CombineSpecs()`.
- **Example 3.3** generates a report incorporating multiple timepoints using the sample `reporting` data saved as part of {gsm}.  

```{r file = system.file("examples", "3_ReportingWorkflow.R", package = "gsm"), eval = FALSE, include = TRUE}

```
