---
title: "v1.9.0 Data Model Update Reference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{v1.9.0 Data Model Update Reference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gsm)
```

# Introduction

gsm v1.9.0 represents a substantial change to the gsm data model. The main changes include new log tables with different names and altered data variables, and new functionality for `Make_Snapshot`. Moving forward with this new model, data will be cumulatively collected and combined to allow for tracking changes through time of a study. Below is a description of the past data for reference along with the new data and highlighted changes in the v1.9.0 release.

------------------------------------------------------------------------

# Data Model pre-v1.9.0

The tables below represent the past data model for a reference

### status_study

| Table        | Column                     | Description                                         |
|------------------|----------------------|--------------------------------|
| status_study | studyid                    | Unique Study ID                                     |
| status_study | enrolled_sites             | Number of enrolled sites                                |
| status_study | enrolled_participants      | Number of enrolled participants                         |
| status_study | planned_sites              | Number of planned sites                                 |
| status_study | planned_participants       | Number of planned participants                          |
| status_study | title                      | Protocol title                                      |
| status_study | nickname                   | Protocol nickname                                   |
| status_study | enrolled_sites_ctms        | Number of enrolled sites from GILDA                     |
| status_study | enrolled_participants_ctms | Number of enrolled participants from GILDA              |
| status_study | fpfv                       | First-patient first visit date                      |
| status_study | lpfv                       | Last-patient first visit date                       |
| status_study | lplv                       | Last-patient last visit date                        |
| status_study | ta                         | Therapeutic Area                                    |
| status_study | indication                 | Indication                                          |
| status_study | phase                      | Phase                                               |
| status_study | status                     | Study Status                                        |
| status_study | rbm_flag                   | Risk-based monitoring flag                          |
| status_study | product                    | Product                                             |
| status_study | protocol_type              | Protocol type                                       |
| status_study | protocol_row_id            | Protocol row ID                                     |
| status_study | est_fpfv                   | Estimated first-patient first visit date from GILDA |
| status_study | est_lpfv                   | Estimated last-patient first visit date from GILDA  |
| status_study | est_lplv                   | Estimated last-patient last visit date from GILDA   |
| status_study | protocol_product_number    | Protocol product number                             |
| status_study | gsm_analysis_date          | Date that snapshot was created                      |

### status_site

| Table       | Column                | Description                       |
|-------------|-----------------------|-----------------------------------|
| status_site | studyid               | Unique Study ID                   |
| status_site | siteid                | Unique Site ID                    |
| status_site | institution           | Institution Name                  |
| status_site | status                | Site Status                       |
| status_site | enrolled_participants | Number of Enrolled Participants   |
| status_site | start_date            | Site Activation Date              |
| status_site | city                  | Site City                         |
| status_site | state                 | Site State                        |
| status_site | country               | Site Country                      |
| status_site | invname               | Investigator Name                 |
| status_site | protocol_row_id       | Protocol row ID                   |
| status_site | site_num              | Site Number                       |
| status_site | site_row_id           | Site row ID                       |
| status_site | pi_number             | Principal Investigator Number     |
| status_site | pi_last_name          | Principal Investigator Last Name  |
| status_site | pi_first_name         | Principal Investigator First Name |
| status_site | is_satellite          | Is site a satellite location      |
| status_site | gsm_analysis_date     | Date that Snapshot was created    |

### status_workflow

| Table           | Column            | Description                                   |
|------------------|---------------------|---------------------------------|
| status_workflow | studyid           | Unique Study ID                               |
| status_workflow | workflowid        | Unique workflow ID                            |
| status_workflow | gsm_version       | Version of GSM used                           |
| status_workflow | active            | Logical indicator if workflow was set to run  |
| status_workflow | status            | Logical indicator if workflow was able to run |
| status_workflow | notes             | Notes on workflow                             |
| status_workflow | gsm_analysis_date | Date that Snapshot was created                |

### status_param

| Table        | Column            | Description                    |
|--------------|-------------------|--------------------------------|
| status_param | studyid           | Unique Study ID                |
| status_param | workflowid        | Unique Workflow ID             |
| status_param | gsm_version       | GSM version of workflow        |
| status_param | param             | Parameter name                 |
| status_param | index             | Index value for parameter      |
| status_param | value             | Value for parameter at index   |
| status_param | gsm_analysis_date | Date that snapshot was created |

### results_summary

| Table           | Column            | Description                                     |
|------------------|----------------------|--------------------------------|
| results_summary | studyid           | Unique Study ID                                 |
| results_summary | workflowid        | Unique Workflow ID                              |
| results_summary | groupid           | Unique Group ID (e.g. SiteID for KRI workflows) |
| results_summary | numerator         | Metric numerator                                |
| results_summary | denominator       | Metric denominator                              |
| results_summary | metric            | Metric value                                    |
| results_summary | score             | Statistical Score                               |
| results_summary | flag              | Flag                                            |
| results_summary | gsm_analysis_date | Date that snapshot was created                  |

### results_analysis

| Table            | Column            | Description                    |
|------------------|-------------------|--------------------------------|
| results_analysis | studyid           | Unique Study ID                |
| results_analysis | workflowid        | Unique Workflow ID             |
| results_analysis | param             | Parameter name                 |
| results_analysis | value             | Value for parameter            |
| results_analysis | gsm_analysis_date | Date that snapshot was created |

### results_bounds

| Table          | Column            | Description                    |
|----------------|-------------------|--------------------------------|
| results_bounds | studyid           | Unique Study ID                |
| results_bounds | workflowid        | Unique Workflow ID             |
| results_bounds | threshold         | Threshold                      |
| results_bounds | numerator         | y value                        |
| results_bounds | denominator       | x value                        |
| results_bounds | log_denominator   | Log X value                    |
| results_bounds | gsm_analysis_date | Date that snapshot was created |

### meta_workflow

| Table         | Column            | Description                            |
|---------------|-------------------|----------------------------------------|
| meta_workflow | workflowid        | Unique Workflow ID                     |
| meta_workflow | gsm_version       | GSM Version of workflow                |
| meta_workflow | group             | Group for workflow - Site              |
| meta_workflow | abbreviation      | Workflow Abbreviation (max 3-letters)  |
| meta_workflow | metric            | Metric Label                           |
| meta_workflow | numerator         | Numerator Label                        |
| meta_workflow | denominator       | Denominator Label                      |
| meta_workflow | outcome           | Metric Outcome Type (e.g. rate)        |
| meta_workflow | model             | Statistical Model used to create score |
| meta_workflow | score             | Score Label                            |
| meta_workflow | data_inputs       | Data inputs for workflow               |
| meta_workflow | data_filters      | Filters applied in workflow            |
| meta_workflow | gsm_analysis_date | Date that snapshot was created         |

### meta_param

| Table      | Column            | Description                          |
|------------|-------------------|--------------------------------------|
| meta_param | workflowid        | Unique Workflow ID                   |
| meta_param | gsm_version       | GSM version of workflow              |
| meta_param | param             | Parameter name                       |
| meta_param | index             | Parameter index                      |
| meta_param | default           | Default value for parameter at index |
| meta_param | configurable      | Is parameter configurable?           |
| meta_param | gsm_analysis_date | Date that snapshot was created       |

------------------------------------------------------------------------

# Data Model post-v1.9.0

The table below represent the new data model

### rpt_site_details
Created using the `status_site` and `results_summary` tables 

| Table            | Column                | Description                         |
|------------------|----------------------|--------------------------------|
| rpt_site_details | studyid               | Unique Study ID                     |
| rpt_site_details | snapshot_date         | Date that snapshot was created      |
| rpt_site_details | siteid                | Unique Site ID                      |
| rpt_site_details | site_num              | Number of site                      |
| rpt_site_details | institution           | Name of site                        |
| rpt_site_details | status                | Status of site                      |
| rpt_site_details | start_date            | Start date of site                  |
| rpt_site_details | invname               | Investigators name                  |
| rpt_site_details | country               | Country site is located             |
| rpt_site_details | state                 | State site is located               |
| rpt_site_details | city                  | City site is located                |
| rpt_site_details | region                | Region of site                      |
| rpt_site_details | enrolled_participants | Enrolled participants               |
| rpt_site_details | planned_participants  | Planned participants                |
| rpt_site_details | num_of_at_risk_kris   | Total number of amber flags per KRI |
| rpt_site_details | num_of_flagged_kris   | Total number of red flags per KRI   |
| rpt_site_details | pt_cycle_id           | Cycle ID                            |
| rpt_site_details | pt_data_dt            | Data ID                             |
| rpt_site_details | gsm_analysis_date     | Date that snapshot was created      |

### rpt_study_details
Created using the `status_study` and `results_summary` tables

| Table             | Column                            | Description                          |
|------------------|----------------------|--------------------------------|
| rpt_study_details | studyid                           | Unique Study ID                      |
| rpt_study_details | snapshot_date                     | Date that snapshot was created       |
| rpt_study_details | title                             | Title of protocol                    |
| rpt_study_details | ta                                | Area of therapy                      |
| rpt_study_details | indication                        | Protocol indication                  |
| rpt_study_details | phase                             | Phase of study                       |
| rpt_study_details | product                           | Name of product                      |
| rpt_study_details | enrolled_sites                    | Sites enrolled in study              |
| rpt_study_details | enrolled_participants             | Participants enrolled in study       |
| rpt_study_details | planned_sites                     | Total planned sites for study        |
| rpt_study_details | planned_participants              | Total planned participants for study |
| rpt_study_details | est_fpfv                          | First patient first visit            |
| rpt_study_details | est_lpfv                          | Last patient first visit             |
| rpt_study_details | est_lplv                          | Last patient last visit              |
| rpt_study_details | status                            | Study status                         |
| rpt_study_details | fpfv                              | First patient first visit            |
| rpt_study_details | lpfv                              | Last patient first visit             |
| rpt_study_details | lplv                              | Last patient last visit              |
| rpt_study_details | study_age                         | Total age of the study in YMD        |
| rpt_study_details | num_of_sites_flagged              | Total number of sites with red flags |
| rpt_study_details | enrolling_sites_with_flagged_kris | Enrolled sites with flagged kris     |
| rpt_study_details | pt_cycle_id                       | Cycle ID                             |
| rpt_study_details | pt_data_dt                        | Data ID                              |
| rpt_study_details | gsm_analysis_date                 | Date that snapshot was created       |

### rpt_qtl_details
Created using the `status_param`, `results_summary`, and `meta_workflow` tables

| Table           | Column            | Description                                      |
|------------------|----------------------|--------------------------------|
| rpt_qtl_details | studyid           | Unique Study ID                                  |
| rpt_qtl_details | snapshot_date     | Date that snapshot was created                   |
| rpt_qtl_details | workflowid        | Unique QTL ID                                    |
| rpt_qtl_details | metric            | QTL name                                         |
| rpt_qtl_details | numerator_name    | Name of the numerator used to calculate metric   |
| rpt_qtl_details | denominator_name  | Name of the denominator used to calculate metric |
| rpt_qtl_details | qtl_value         | QTL numerical metric                             |
| rpt_qtl_details | base_metric       | Name of numerator over name of denominator       |
| rpt_qtl_details | numerator_value   | numerator numerical metric                       |
| rpt_qtl_details | denominator_value | denominator numerical metric                     |
| rpt_qtl_details | qtl_score         | QTL statistical score                            |
| rpt_qtl_details | flag              | QTL flag value                                   |
| rpt_qtl_details | threshold         | Thershold used to determine score                |
| rpt_qtl_details | abbreviation      | Abbreviated QTL name                             |
| rpt_qtl_details | outcome           | Units representing outcome                       |
| rpt_qtl_details | model             | Model used for evaluation                        |
| rpt_qtl_details | meta_score        | Statistical score used                           |
| rpt_qtl_details | data_inputs       | Data used as inputs for the model                |
| rpt_qtl_details | data_filters      | Filters applied to data inputs                   |
| rpt_qtl_details | gsm_version       | gsm version used for analysis                    |
| rpt_qtl_details | group             | The grouping used for the analysis               |
| rpt_qtl_details | pt_cycle_id       | Cycle ID                                         |
| rpt_qtl_details | pt_data_dt        | Data ID                                          |
| rpt_qtl_details | gsm_analysis_date | Date that snapshot was created                   |

### rpt_kri_details
Created using the `meta_workflow`, `status_site`, `status_workflow` and `results_summary` tables

| Table           | Column               | Description                                          |
|------------------|----------------------|--------------------------------|
| rpt_kri_details | studyid              | Name of study                                        |
| rpt_kri_details | snapshot_date        | Date that snapshot was created                       |
| rpt_kri_details | workflowid           | Unique KRI                                           |
| rpt_kri_details | metric               | Name of the KRI                                      |
| rpt_kri_details | abbreviation         | KRI abbreviation                                     |
| rpt_kri_details | kri_description      | KRI Description                                      |
| rpt_kri_details | base_metric          | Numerator over denominator                           |
| rpt_kri_details | numerator            | Meta workflow numerator                              |
| rpt_kri_details | denominator          | Meta workflow denominator                            |
| rpt_kri_details | num_of_sites_at_risk | Total number of grouped sites with amber flags       |
| rpt_kri_details | num_of_sites_flagged | Total number of grouped sites with red flags         |
| rpt_kri_details | outcome              | The units output in the outcome of the KRI           |
| rpt_kri_details | model                | The method of calculating the score for the KRI      |
| rpt_kri_details | score                | The type of score used for the KRI                   |
| rpt_kri_details | data_inputs          | Data used as inputs                                  |
| rpt_kri_details | data_filters         | Data filtering parameters                            |
| rpt_kri_details | gsm_version          | version of gsm used for analysis                     |
| rpt_kri_details | group                | group targeted by KRI                                |
| rpt_kri_details | total_num_of_sites   | Total number of sites in study                       |
| rpt_kri_details | pt_cycle_id          | Cycle ID                                             |
| rpt_kri_details | pt_data_dt           | Data date                                            |
| rpt_kri_details | active               | Logical value indicating if kri is currently active  |
| rpt_kri_details | status               | Logical value indicating if kri was successfully run |
| rpt_kri_details | notes                | User input notes per kri                             |
| rpt_kri_details | gsm_analysis_date    | Date that snapshot was created                       |

### rpt_site_kri_details
Created using the `meta_workflow`, `status_param`, `meta_param`, and `results_summary` tables

| Table                | Column                  | Description                                                    |
|------------------|----------------------|--------------------------------|
| rpt_site_kri_details | studyid                 | Unique Study ID                                                |
| rpt_site_kri_details | snapshot_date           | Date that snapshot was created                                 |
| rpt_site_kri_details | siteid                  | Unique Site ID                                                 |
| rpt_site_kri_details | workflowid              | Unique KRI ID                                                  |
| rpt_site_kri_details | metric_value            | KRI value                                                      |
| rpt_site_kri_details | score                   | KRI score based on value                                       |
| rpt_site_kri_details | numerator_value         | Numerator used to calculate metric                             |
| rpt_site_kri_details | denominator_value       | Denominator used to calculate metric                           |
| rpt_site_kri_details | flag_value              | Flag value reported based on score                             |
| rpt_site_kri_details | no_of_consecutive_loads | The number of times the site KRI has been continuously flagged |
| rpt_site_kri_details | upper_threshold         | Value from status_param when index = 3                         |
| rpt_site_kri_details | lower_threshold         | Value from status_param when index = 2                         |
| rpt_site_kri_details | bottom_lower_threshold  | Value from status_param when index = 1                         |
| rpt_site_kri_details | top_upper_threshold     | Value from status_param when index = 4                         |
| rpt_site_kri_details | metric                  | Name of KRI                                                    |
| rpt_site_kri_details | country_aggregate       | Country aggregate                                              |
| rpt_site_kri_details | study_aggregate         | Study aggregate                                                |
| rpt_site_kri_details | numerator_name          | Description of Numerator                                       |
| rpt_site_kri_details | denominator_name        | Description of Denominator                                     |
| rpt_site_kri_details | pt_cycle_id             | Cycle ID                                                       |
| rpt_site_kri_details | pt_data_dt              | Data ID                                                        |
| rpt_site_kri_details | gsm_analysis_date       | Date that snapshot was created                                 |

### rpt_kri_bounds_details
Created using the `results_bounds` table

| Table                  | Column            | Description                          |
|---------------------|---------------------|------------------------------|
| rpt_kri_bounds_details | studyid           | Unique Study ID                      |
| rpt_kri_bounds_details | snapshot_date     | Date that snapshot was created       |
| rpt_kri_bounds_details | workflowid        | Unique KRI ID                        |
| rpt_kri_bounds_details | threshold         | KRI value                            |
| rpt_kri_bounds_details | numerator         | Threshold                            |
| rpt_kri_bounds_details | denominator       | Numerator used to calculate metric   |
| rpt_kri_bounds_details | log_denominator   | Denominator used to calculate metric |
| rpt_kri_bounds_details | pt_cycle_id       | Log transformed denominator          |
| rpt_kri_bounds_details | pt_data_dt        | Cycle ID                             |
| rpt_kri_bounds_details | gsm_analysis_date | Data ID                              |

### rpt_qtl_threshold_param
Created using the `meta_param` and `status_param` tables

| Table                   | Column            | Description                                             |
|---------------------|---------------------|------------------------------|
| rpt_qtl_threshold_param | studyid           | Unique Study ID                                         |
| rpt_qtl_threshold_param | snapshot_date     | Date that snapshot was created                          |
| rpt_qtl_threshold_param | workflowid        | Unique QTL ID                                           |
| rpt_qtl_threshold_param | gsm_version       | version of gsm used for analysis                        |
| rpt_qtl_threshold_param | param             | Parameter used for QTL                                  |
| rpt_qtl_threshold_param | index             | Index used for threshold values                         |
| rpt_qtl_threshold_param | default_s         | Default values for parameters                           |
| rpt_qtl_threshold_param | configurable      | Logical field indicating if parameter can be configured |
| rpt_qtl_threshold_param | pt_cycle_id       | Cycle ID                                                |
| rpt_qtl_threshold_param | pt_data_dt        | Data ID                                                 |
| rpt_qtl_threshold_param | gsm_analysis_date | Date that snapshot was created                          |

### rpt_kri_threshold_param
Created using the `meta_param` and `status_param` tables

| Table                   | Column            | Description                                             |
|---------------------|---------------------|------------------------------|
| rpt_kri_threshold_param | studyid           | Unique Study ID                                         |
| rpt_kri_threshold_param | snapshot_date     | Date that snapshot was created                          |
| rpt_kri_threshold_param | workflowid        | Unique KRI ID                                           |
| rpt_kri_threshold_param | gsm_version       | version of gsm used for analysis                        |
| rpt_kri_threshold_param | param             | Parameter used for QTL                                  |
| rpt_kri_threshold_param | index             | Index used for threshold values                         |
| rpt_kri_threshold_param | default_s         | Default values for parameters                           |
| rpt_kri_threshold_param | configurable      | Logical field indicating if parameter can be configured |
| rpt_kri_threshold_param | pt_cycle_id       | Cycle ID                                                |
| rpt_kri_threshold_param | pt_data_dt        | Data ID                                                 |
| rpt_kri_threshold_param | gsm_analysis_date | Date that snapshot was created                          |

### rpt_qtl_analysis
Created using the `results_analysis`` table

| Table            | Column            | Description                    |
|------------------|-------------------|--------------------------------|
| rpt_qtl_analysis | studyid           | Unique Study ID                |
| rpt_qtl_analysis | snapshot_date     | Date that snapshot was created |
| rpt_qtl_analysis | workflowid        | Unique QTL ID                  |
| rpt_qtl_analysis | param             | Parameter used for QTL         |
| rpt_qtl_analysis | value             | Score for the QTL              |
| rpt_qtl_analysis | pt_cycle_id       | Cycle ID                       |
| rpt_qtl_analysis | pt_data_dt        | Data ID                        |
| rpt_qtl_analysis | gsm_analysis_date | Date that snapshot was created |

------------------------------------------------------------------------

# Highlight Changes

## Dropped Variables

| Table            | Column            | Description                    |
|------------------|-------------------|--------------------------------|
| status_site | protocol_row_id       | Protocol row ID                   |
| status_site | site_row_id           | Site row ID                       |
| status_site | pi_number             | Principal Investigator Number     |
| status_site | pi_last_name          | Principal Investigator Last Name  |
| status_site | pi_first_name         | Principal Investigator First Name |
| status_site | is_satellite          | Is site a satellite location      |
| status_study | nickname                   | Protocol nickname                                   |
| status_study | enrolled_sites_ctms        | Number of enrolled sites from GILDA                     |
| status_study | enrolled_participants_ctms | Number of enrolled participants from GILDA              |
| status_study | rbm_flag                   | Risk-based monitoring flag                          |
| status_study | protocol_type              | Protocol type                                       |
| status_study | protocol_row_id            | Protocol row ID                                     |
| status_param | value             | Value for parameter at index   |

## Added Variables

| Table            | Column            | Description                    |
|------------------|-------------------|--------------------------------|
| All new tables | snapshot_date                     | Date that snapshot was created       |
| All new tables | pt_cycle_id           | Cycle ID                            |
| All new tables | pt_data_dt            | Data date                             |
| rpt_site_details | region                | Region of site                      |
| rpt_site_details | planned_participants  | Planned participants                |
| rpt_study_details | study_age                         | Total age of the study in YMD        |
| rpt_study_details | enrolling_sites_with_flagged_kris | Enrolled sites with flagged kris     |
| rpt_qtl_details | base_metric       | Name of numerator over name of denominator       |
| rpt_site_kri_details | no_of_consecutive_loads | The number of times the site KRI has been continuously flagged |
| rpt_site_kri_details | country_aggregate       | Country aggregate                                              |
| rpt_site_kri_details | study_aggregate         | Study aggregate                                                |

## New Functionality
Previously `Make_Snapshot` was functionally used to extrapolate meaningful statistics and information for a study at a single moment in time. These snapshots were then intended to be saved in a location using `Save_Snapshot` and if longitudinal data was desired the user would then call `Augment_Snapshot`to combine all past snapshot outputs into a longitudinal snapshot.

The new functionality eliminates the need for `Augment_Snapshot` (removed in the v1.9.0 release). `Make_Snapshot` will now have an `lPrevSnapshot` argument for the previous snapshot object (snapshot.rds file), this is designed to accumulate and combine data as snapshots are run. 

## `UpdateSnapshotLogs`
For backwards compatibility a new funtion called `UpdateSnapshotLogs` is used to retrofit the past saved snapshots, that were saved prior to v1.9.0, to the new data model. This function will convert the `lSnapshot` and `lStackedSnapshots` logs to the new data model logs (rpt_* tables)
